---
title: "Iris Data"
output:
  pdf_document: default
  word_document: default
date: "May 2, 2019"
---

```{r setup, include=true}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Load Data and Libraries
```{r}
data(iris)
library(ggplot2)
library(gridExtra)
library(ggplot2)
library(cowplot)
library(dplyr)
library(grid)
library(ICSNP)
library(VGAM)
library(car)
library(openxlsx)
# setwd() set the work directionary accordingly

SL=iris$Sepal.Length
SW=iris$Sepal.Width
PL=iris$Petal.Length
PW=iris$Petal.Width
Species=iris$Species
```

# 1. Exploratory Analysis
```{r}
head(iris)
data_sample = some(iris)
# write the result in excel file
# write.xlsx(data_sample, 'irisdata.xlsx') 
```

## 1.1 Boxplots
```{r}
irisVer <- subset(iris, Species == "versicolor")
irisSet <- subset(iris, Species == "setosa")
irisVir <- subset(iris, Species == "virginica")
par(mfrow=c(1,3),mar=c(6,3,2,1))
boxplot(irisSet[,1:4], main="Setosa",ylim = c(0,8),las=2)
boxplot(irisVer[,1:4], main="Versicolor",ylim = c(0,8),las=2)
boxplot(irisVir[,1:4], main="Virginica",ylim = c(0,8),las=2)
```

## 1.2 Histograms

### 1.2.1 Sepal. Length
```{r}
par(mfrow=c(1,3))
hist(irisVer$Sepal.Length,xlim=c(0,8),ylim=c(0,20))
hist(irisSet$Sepal.Length,xlim=c(0,8),ylim=c(0,20))
hist(irisVir$Sepal.Length,xlim=c(0,8),ylim=c(0,20))
```

### 1.2.2 Sepal.Width
```{r}
par(mfrow=c(1,3))
hist(irisVer$Sepal.Width,xlim=c(0,8),ylim=c(0,20))
hist(irisSet$Sepal.Width,xlim=c(0,8),ylim=c(0,20))
hist(irisVir$Sepal.Width,xlim=c(0,8),ylim=c(0,20))
```

### 1.2.3 Petal.Length
```{r}
par(mfrow=c(1,3))
hist(irisVer$Petal.Length,xlim=c(0,8),ylim=c(0,20))
hist(irisSet$Petal.Length,xlim=c(0,8),ylim=c(0,20))
hist(irisVir$Petal.Length,xlim=c(0,8),ylim=c(0,20))
```

### 1.2.4 Petal.Width
```{r}
par(mfrow=c(1,3))
hist(irisVer$Petal.Length,xlim=c(0,8),ylim=c(0,20))
hist(irisSet$Petal.Length,xlim=c(0,8),ylim=c(0,20))
hist(irisVir$Petal.Length,xlim=c(0,8),ylim=c(0,20))
```

## 1.3 Scatterplots
```{r}
# If we want to find out which characteristics are most distinguishing between iris plants, we have to make many individual plots and hope we can see distinguishing patterns:
p1 <- ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, color=Species)) + geom_point()
p2 <- ggplot(iris, aes(x=Petal.Length, y=Petal.Width, color=Species)) + geom_point()
p3 <- ggplot(iris, aes(x=Sepal.Length, y=Petal.Length, color=Species)) + geom_point()
p4 <- ggplot(iris, aes(x=Sepal.Width, y=Petal.Width, color=Species)) + geom_point()
plot_grid(p1, p2, p3, p4, labels = "AUTO")
```


# 2. Covariance-based Principal Components Analysis
## 2.1 Preparation and Calculation
```{r}
# Treating y = (SL, SW, PL, PW)> as the response vector, and ignoring the Species factor for the moment, perform a covariance-based PCA.

# The basic steps in PCA are to (i) prepare a data frame that holds only the numerical columns of interest, (ii) scale the data to 0 mean and unit variance, and (iii) do the PCA with the function prcomp()

iris %>% select(-Species) %>% # remove Species column
  scale(FALSE) %>%            # covariance-based
  prcomp(TRUE) ->            # do PCA ()
  pca_cov                     # store result as `pca_cov`

# now display the results from the PCA analysis
pca_cov
names(pca_cov)

# save the result in excel
# write.xlsx(pca_cov,"pca_cov.xlsx")
Eigenvalues_cov <- eigen(cov(iris[1:4]))
Eigenvalues_cov
```

## 2.2 Scatterplots of PC Scores.
```{r}
#Comment on the plots. Do they all appear to exhibit random scatter? Or do any suggest the presence of structure? Discuss.
pca_cov_data <- data.frame(pca_cov$x, Species=iris$Species)
pairs(pca_cov_data[1:4])
```

## 2.3 Scree Plot
```{r}
# discuss how many PCs it would be sensible to retain.
screeplot(pca_cov,type = "line")
```

# 2.4 PC Scores Comparison
Describe the resulting plot in words. Obtain scatterplots for other pairs of PC scores. Discuss and summaries your findings.
```{r}
# plot the PC scores associated with the PC1 against the PC2 using the command

p1 <- ggplot(pca_cov_data, aes(x=PC1, y=PC2, color=Species)) + geom_point()
p2 <- ggplot(pca_cov_data, aes(x=PC1, y=PC3, color=Species)) + geom_point()
p3 <- ggplot(pca_cov_data, aes(x=PC1, y=PC4, color=Species)) + geom_point()
p4 <- ggplot(pca_cov_data, aes(x=PC2, y=PC3, color=Species)) + geom_point()
p5 <- ggplot(pca_cov_data, aes(x=PC2, y=PC4, color=Species)) + geom_point()
p6 <- ggplot(pca_cov_data, aes(x=PC3, y=PC4, color=Species)) + geom_point()
plot_grid(p1, p2, p3, p4, p5,p6, nrow = 3)
```

# 3. Correlation-based PCA
Now perform a correlation-based PCA, performing similar calculations to those indicated in Sections 2.

## 3.1 Preparation and Calculation
```{r}
# The basic steps in PCA are to (i) prepare a data frame that holds only the numerical columns of interest, (ii) scale the data to 0 mean and unit variance, and (iii) do the PCA with the function prcomp()

iris %>% select(-Species) %>% # remove Species column
  scale(TRUE) %>%             # correlation-based (TRUE)
  prcomp() ->                 # do PCA 
  pca_cor                     # store result as `pca_cor`

# now display the results from the PCA analysis
pca_cor

# save the result in excel
# write.xlsx(pca_cor,"pca_cor.xlsx")

Eigenvalues_cor <- eigen(cor(iris[1:4]))
Eigenvalues_cor
```
## 3.2 Scatterplots of PC Scores.
```{r}
#Comment on the plots. Do they all appear to exhibit random scatter? Or do any suggest the presence of structure? Discuss.
pca_cor_data <- data.frame(pca_cor$x, Species=iris$Species)
pairs(pca_cor_data[1:4])
```

## 3.3 Scree Plot
```{r}
# discuss how many PCs it would be sensible to retain.
screeplot(pca_cor, type = "line")
```

# 3.4 PC Scores Comparison
Describe the resulting plot in words. Obtain scatterplots for other pairs of PC scores. Discuss and summaries your findings.
```{r}
# plot the PC scores associated with the PC1 against the PC2 using the command
pp1 <- ggplot(pca_cor_data, aes(x=PC1, y=PC2, color=Species)) + geom_point()
pp2 <- ggplot(pca_cor_data, aes(x=PC1, y=PC3, color=Species)) + geom_point()
pp3 <- ggplot(pca_cor_data, aes(x=PC1, y=PC4, color=Species)) + geom_point()
pp4 <- ggplot(pca_cor_data, aes(x=PC2, y=PC3, color=Species)) + geom_point()
pp5 <- ggplot(pca_cor_data, aes(x=PC2, y=PC4, color=Species)) + geom_point()
pp6 <- ggplot(pca_cor_data, aes(x=PC3, y=PC4, color=Species)) + geom_point()
plot_grid(pp1, pp2, pp3, pp4, pp5,pp6, nrow = 3)
```

# 4. 
```{r}
# Discuss the similarities and differences of the covariance-based and correlation-based approaches for performing PCA on the Iris dataset. Which of these two approaches, if either, do you think is preference for the Iris dataset?
```

# 5. Hotelling's T-square Statistic
```{r}
# Compare the species pairwise using Hotelling's T2 statistic.
irisSet=irisSet[1:4]
irisVer=irisVer[1:4]
irisVir=irisVir[1:4]
HotellingsT2(irisSet,irisVer)
HotellingsT2(irisSet,irisVir)
HotellingsT2(irisVir,irisVer)
```

# 6. Multivariate Linear Regression
```{r}
mod.iris <- lm(cbind(Sepal.Length, Sepal.Width, Petal.Length, Petal.Width)~ Species, data=iris)
summary(mod.iris)
```

