---
title: "Credit Limit MDP"
author: "Joey Li"
date: "July 31, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Data Preparation
```{r}
rm(list = ls())
library(tidyverse)  
library(cluster)
library(markovchain)

credit <- read.csv("fidi_carte.csv",header = TRUE)
credit <- arrange(credit, X)
credit <- credit[(1:(nrow(credit)/2)),-(11:13)]
colnames(credit) <- c('No.', 'KEY', 'DATE', 'SCORE', 'STATUS', 'TIME', 'RATING_BEFORE', 'PD_BEFORE', 'BUREAU_SCORE', 'PLAFOND')
```

# 2. Regroup Status According to PLAFOND
## 2.1 Use K Means to Decide Groups
```{r}
## K Means will find out the best groups according to the data characteristics, instead of manual definition.
## Where centers = 6, indicates we want the data be devided into 6 groups.
## And then I store the group number from KNN in the table as LIMITGROUP. 
LGroup <- kmeans(credit$PLAFOND, centers = 6, nstart = 50)
print(LGroup)
credit$LIMITGROUP <- LGroup$cluster
```

## 2.2 Count Numbers of Good/Bad Under in Limit Group
```{r}
## To display the summary table of limit group and status, just use table() funcion.
table_limit <- credit[,c(5,11)]
table_limit <- table(table_limit$LIMITGROUP, table_limit$STATUS)
table_limit
```

# 3. Regroup Status According to SCORE
## 3.1 Use K-Means Method to Decide Groups
```{r}
## Same as before, I use K means to define groups again, by using centers = 10
## And then I store the results as SCOREGROUP. 
SGroup <- kmeans(credit$SCORE, centers = 10, nstart = 50)
print(SGroup)
credit$SCOREGROUP <- SGroup$cluster
```

## 3.2 Count Numbers of Good/Bad in Each Score Group 
```{r}
table_score <- credit[,c(5,12)]
table_score <- table(table_score$SCOREGROUP, table_score$STATUS)
table_score
```

# 4. Recongize Default Loss and Output Loss Matrix
```{r}
## if default (STATUS == '1), the bank loses the amount of PLAFOND
## Just use the simple logic clause and assign values directly.
credit$LOSS[credit$STATUS == '1'] <- -credit$PLAFOND

## Write table as Loss Matrix
LOSSMATRIX <- with(credit, tapply(LOSS, list(SCOREGROUP, LIMITGROUP), FUN = sum))

colnames(LOSSMATRIX) <- c("Limit1","Limit2","Limit3","Limit4","Limit5","Limit6")
rownames(LOSSMATRIX) <- c("Score1","Score2","Score3","Score4","Score5","Score6","Score7","Score8","Score9","Score10")

data.frame(LOSSMATRIX)
```

# 5. Prepare Transition Matrix
```{r}


```

# 6. Try Supervised Marchine Learning Techniques
```{r}
library(rpart)
colnames(credit)

credit_try <- glm(STATUS ~ LIMITGROUP + SCOREGROUP + SCORE + TIME + PD_BEFORE + PLAFOND, 
                  family = 'binomial', credit)
summary(credit_try)

fit <- rpart(STATUS ~ LIMITGROUP + SCOREGROUP + SCORE + TIME + RATING_BEFORE + PD_BEFORE + PLAFOND, credit)
rplot(fit)

```







